---
title: "Random forest acoustic space budgies 2018"
author: <a href="http://marceloarayasalas.weebly.com/">Marcelo Araya-Salas</a>
 &nbsp; 
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: kable
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: no
editor_options: 
  chunk_output_type: console
---

```{r packages, echo=F, eval=T, message=FALSE, warning=FALSE }

rm(list = ls())

# unload all non-based packages
out <- sapply(paste('package:', names(sessionInfo()$otherPkgs), sep = ""), function(x) try(detach(x, unload = FALSE, character.only = TRUE), silent = T))

x<-c("warbleR", "readxl", "RColorBrewer", "ggplot2", "parallel", "ade4", "fossil", "vegan", "AICcmodavg", "nlme", "ecodist", "lme4", "Hmisc",  "compiler", "smacof", "pbapply", "caret", "randomForest", "corrplot", "kableExtra", "adehabitatHR", "spatstat", "raster", "cowplot", "MuMIn", "sjstats", "lmerTest")

aa <- lapply(x, function(y) {
  if(!y %in% installed.packages()[,"Package"])  {if(!y %in% c("warbleR", "Rraven")) install.packages(y) else devtools::install_github(paste0("maRce10/", y))
}
try(require(y, character.only = T), silent = T)
  })

```

```{r functions and parameters, eval = T, echo = F}

options("digits" = 6, "digits.secs" = 5, knitr.table.format = "html") 

cols <- brewer.pal(10,"Spectral")

prll <- detectCores() -1

range.mat.0.1 <- function(mat, verbose = F){
  
if(verbose)  
print(paste( "original range:", paste(round(range(mat[upper.tri(mat, diag=FALSE)]), 2), collapse = "-")))

new.values <- mat[upper.tri(mat, diag=FALSE)] -  min(mat[upper.tri(mat, diag=FALSE)])  

mat[upper.tri(mat)] <- new.values

b <- mat[upper.tri(mat, diag=FALSE)] <- mat[upper.tri(mat, diag=FALSE)]/  max(mat[upper.tri(mat, diag=FALSE)])  

mat <- t(mat)

mat[upper.tri(mat, diag=FALSE)] <- b

if(verbose)  
print(paste( "new range:", paste(round(range(mat[upper.tri(mat, diag=FALSE)]), 2), collapse = "-")))

return(mat)
}

range.mat.0.1 <- cmpfun(range.mat.0.1)

path.recs <- "/media/twright/DATA/consolidated_files/"

source("~/Dropbox/Documentos_R/Myfunctions/myfunctions.R")

plot_acu_dens <- function(Y, x, mds_prox, ovlp)
{   
     Ypre <- Y[Y$period == "pre", ]
  Ypost <- Y[Y$period == "post", ]
  Gpost <- mds_prox[mds_prox$period == "post" & mds_prox$Cage == Y$Cage[1] & mds_prox$Bird.Name != bnm[x], ]
Gpre <- mds_prox[mds_prox$period == "pre" & mds_prox$Cage == Ypre$Cage[1] & mds_prox$Bird.Name != bnm[x], ]  
Y <- rbind(Y, Gpost)
  
  #select calls closest to mean acoustic space
 if(nrow(Ypre) > 0)
   {
   mt <- rbind(data.frame(MDS1 = mean(Ypre[,"MDS1"]),MDS2 = mean(Ypre[,"MDS2"])), Ypre[,c("MDS1", "MDS2")])
   
   fspre <- Ypre$sound.files[which.min(as.matrix(dist(mt, upper = T))[-1,1])]
        Zpre <- mtdt[mtdt$sound.files == fspre,, drop = FALSE][1,]
         predur <- Zpre$end - Zpre$start
         
         } else predur <- 0
        
  if(nrow(Ypost) > 0)
{    
   mt2 <- rbind(data.frame(MDS1 = mean(Ypost[,"MDS1"]),MDS2 = mean(Ypost[,"MDS2"])), Ypost[,c("MDS1", "MDS2")])
    fspost <- Ypost$sound.files[which.min(as.matrix(dist(mt2, upper = T))[-1, 1])]
    Zpost <- mtdt[mtdt$sound.files == fspost, , drop = FALSE][1,]
    postdur <- Zpost$end - Zpost$start
 } else postdur <- 0
   
  if(nrow(Gpost) > 0)
{
         mt3 <- rbind(data.frame(MDS1 = mean(Gpost[,"MDS1"]),MDS2 = mean(Gpost[,"MDS2"])), Gpost[,c("MDS1", "MDS2")])
    fspost2 <- Gpost$sound.files[which.min(as.matrix(dist(mt3, upper = T))[-1, 1])]
    Zpost2 <- mtdt[mtdt$sound.files == fspost2, , drop = FALSE][1, , drop = FALSE]
    postdur2 <- Zpost2$end - Zpost2$start
}  else postdur2 <- 0


  if(nrow(Gpre) > 0)
{
         mt4 <- rbind(data.frame(MDS1 = mean(Gpre[,"MDS1"]),MDS2 = mean(Gpre[,"MDS2"])), Gpre[,c("MDS1", "MDS2")])
    fspre2 <- Gpre$sound.files[which.min(as.matrix(dist(mt4, upper = T))[-1, 1])]
}

  #adjust all to the same duration
maxdur <- max(postdur, predur, postdur2)
    if(maxdur > predur & nrow(Ypre) > 0) 
    {
    Zpre$start <- (Zpre$start - (maxdur-predur)/2)
  Zpre$end <- (Zpre$end + (maxdur-predur)/2)
  } 

  if(maxdur > postdur & nrow(Ypost) > 0) 
    {
    Zpost$start <- (Zpost$start - (maxdur-postdur)/2)
  Zpost$end <- (Zpost$end + (maxdur-postdur)/2)
  } 
  
  if(maxdur > postdur2 & exists("Zpost2"))
    if(nrow(Zpost2) > 0)
    {
    Zpost2$start <- (Zpost2$start - (maxdur-postdur)/2)
  Zpost2$end <- (Zpost2$end + (maxdur-postdur)/2)
  } 

   if(nrow(Ypre) > 0)   
    wpre <- read_wave(mtdt, index = which(mtdt$sound.files == Zpre$sound.files), from = ifelse(Zpre$start < 0, 0, Zpre$start), to = Zpre$end)

if(nrow(Ypost) > 0)
      wpost <- read_wave(mtdt, index = which(mtdt$sound.files == Zpost$sound.files), from = Zpost$start, to = Zpost$end)

if(exists("Zpost2"))
if(nrow(Zpost2) > 0)
        wpost2 <- read_wave(mtdt, index = which(mtdt$sound.files == Zpost2$sound.files), from = Zpost2$start, to = Zpost2$end)

  nf <- layout(matrix(c(9, 9, 9, 9, 1, 2, 5, 8, 1, 3, 6, 8, 1, 4, 7, 8), nr = 4, byrow=TRUE), widths=c(0.2, 1.5, 0.11, 3), heights=c(0.2, 1, 1, 1.4))
par(mar = c(0, 0, 0, 0))
plot(1, col = "white", frame.plot = F)
 text(x = 1, y = 1.05, "Frequency (kHz)", srt = 90, cex = 2) 
  par(mar = c(0, 2, 0, 0))
  
  if(nrow(Ypre) > 0)  
  spectro(wpre, f = wpre@samp.rate, wl = wl, ovlp = ovlp, scale = F, osc = F, flim = c(1.1, 5.9), palette = reverse.gray.colors.1, grid = F, axisX = F, tlab = "", cexaxis = 2) else
  plot(1, col = "white", frame.plot = F, axes = F)

if(nrow(Ypost) > 0)  
  spectro(wpost, f = wpost@samp.rate, wl = wl, ovlp = ovlp, scale = F, osc = F, flim = c(1.1, 5.9), palette = reverse.gray.colors.1, grid = F,axisX = F, tlab = "", cexaxis = 2) else
  plot(1, col = "white", frame.plot = F, axes = F)
 
if(exists("Zpost2"))
{if(nrow(Zpost2) > 0)  
{
  par(mar = c(4, 2, 0, 0))  
  spectro(wpost2, f = wpost2@samp.rate, wl = wl, ovlp = ovlp, scale = F, osc = F, flim = c(1.1, 5.9), palette = reverse.gray.colors.1, grid = F, cexaxis = 2, cexlab = 2)
}
  } else  plot(1, col = "white", frame.plot = F, axes = F)

alpha <- 0.5
colpre <- adjustcolor( "red2", alpha.f = alpha)  
colpost <- adjustcolor( "cyan4", alpha.f = alpha)  
colgrp <- adjustcolor( "gray", alpha.f = alpha)  

par(mar = c(0, 0, 0, 0))
plot(1, col = "white", frame.plot = F, axes = F)
rect(par("usr")[1],par("usr")[3],par("usr")[2],par("usr")[4],col = colpre)

par(mar = c(0, 0, 0, 0))
plot(1, col = "white", frame.plot = F, axes = F)
rect(par("usr")[1],par("usr")[3],par("usr")[2],par("usr")[4],col = colpost)

par(mar = c(4, 0, 0, 0))
plot(1, col = "white", frame.plot = F, axes = F, xlab = "")
rect(par("usr")[1],par("usr")[3],par("usr")[2],par("usr")[4],col = colgrp)

par(mar = c(4, 6, 0, 0))  
basecex <- 2.3

xlim <- range(acou.space[, 1])
xlim[2] <-  xlim[2] * 1.15

ylim <- range(acou.space[, 2])
ylim[2] <-  ylim[2] * 1.45

plot(x = Y$MDS1, y = Y$MDS2, col = "white", pch = 20, cex = basecex,  xlab = "MDS 1", ylab= "MDS 2", xlim = xlim, ylim = ylim, cex.lab = 2) 

rect(par("usr")[1],par("usr")[3],par("usr")[2],par("usr")[4], col = adjustcolor(cols[7], 0.1))


par(bg = "white")
polygon(x = acou.space[,1], y = acou.space[,2], col = "white", border = "white", lwd =1)

usr <-  par("usr")

if(nrow(Ypre) > 0) 
{
pp.coords.pre <- as.ppp(as.matrix(Ypre[, c("MDS1", "MDS2")]), usr)
r.coords.pre <- raster(density.ppp(pp.coords.pre))

msk.post <- mask(r.coords.pre, acu.sp.sp)

image(msk.post, add = T, col=palpre)

}


if(nrow(Ypost) > 0)  
{  

pp.coords.post <- as.ppp(as.matrix(Ypost[, c("MDS1", "MDS2")]), usr)
r.coords.post <- raster(density.ppp(pp.coords.post))

msk.post <- mask(r.coords.post, acu.sp.sp)

image(msk.post, add = T, col=palpost)
  }

polygon(x = acou.space[,1], y = acou.space[,2], col = adjustcolor("white", 0), border =  "white", lwd =4)
polygon(x = acou.space[,1], y = acou.space[,2], col = adjustcolor("white", 0), border =  adjustcolor(cols[7], 0.1), lwd =4)

legend(usr[2]-((usr[2] - usr[1]) * 0.45), usr[4]-((usr[4] - usr[3]) * 0.002), lty=c(1,1), col=c(colpre, colpost, colgrp, "black"), legend = c("Pre", "Post", "Pre group centroid", "Post group centroid"), pch = 20, lwd = 0, pt.cex= basecex * 1.5, cex = basecex * 0.5)

#post group centroid
if(nrow(Gpost) > 0) 
{
    points(x = Y$MDS1[Y$sound.files == fspost2], y = Y$MDS2[Y$sound.files == fspost2], col = "black", pch = 20, cex = basecex)
    }

# pre group centroid
if(nrow(Gpre) > 0) 
{
    points(x = Gpre$MDS1[Gpre$sound.files == fspre2], y = Gpre$MDS2[Gpre$sound.files == fspre2], col = "gray", pch = 20, cex = basecex * 1.15)
}

# individual pre post centroids
if(nrow(Ypre) > 0) 
{points(x = Y$MDS1[Y$sound.files == fspre], y = Y$MDS2[Y$sound.files == fspre], col = "black", pch = 20, cex = basecex * 1.4)
points(x = Y$MDS1[Y$sound.files == fspre], y = Y$MDS2[Y$sound.files == fspre], col = "white", pch = 20, cex = basecex)  
points(x = Y$MDS1[Y$sound.files == fspre], y = Y$MDS2[Y$sound.files == fspre], col = colpre, pch = 20, cex = basecex)
}

if(nrow(Ypost) > 0) 
{   points(x = Y$MDS1[Y$sound.files == fspost], y = Y$MDS2[Y$sound.files == fspost], col = "black", pch = 20, cex = basecex * 1.4)
  points(x = Y$MDS1[Y$sound.files == fspost], y = Y$MDS2[Y$sound.files == fspost], col = "white", pch = 20, cex = basecex)
    points(x = Y$MDS1[Y$sound.files == fspost], y = Y$MDS2[Y$sound.files == fspost], col = colpost, pch = 20, cex = basecex)
}


par(mar = c(0, 0, 0, 0))
plot(1, col = "white", frame.plot = F, axes = F)
text(1, labels = paste0(bnm[x], " (", Y$Treatment[1], ") post.ovlp =", ovlp), cex = 2)
# dev.off()
}

gif_acu_dens <- function(Y, stp = 20)
{   
  Y <- Y[order(Y$date), ]
  Ypre <- Y[Y$period == "pre", ]
  Ypost <- Y[Y$period == "post", ]

  #select calls closest to mean acoustic space
 if(nrow(Ypre) > 0)
   {
   mt <- rbind(data.frame(MDS1 = mean(Ypre[,"MDS1"]),MDS2 = mean(Ypre[,"MDS2"])), Ypre[,c("MDS1", "MDS2")])
   
   fspre <- Ypre$sound.files[which.min(as.matrix(dist(mt, upper = T))[-1,1])]
        Zpre <- mtdt[mtdt$sound.files == fspre, , drop = FALSE][1,]
         predur <- Zpre$end - Zpre$start
         
         } else predur <- 0
        
  if(nrow(Ypost) > 0)
{    
   mt2 <- rbind(data.frame(MDS1 = mean(Ypost[,"MDS1"]),MDS2 = mean(Ypost[,"MDS2"])), Ypost[,c("MDS1", "MDS2")])
    fspost <- Ypost$sound.files[which.min(as.matrix(dist(mt2, upper = T))[-1, 1])]
    Zpost <- mtdt[mtdt$sound.files == fspost, , drop = FALSE][1,]
    postdur <- Zpost$end - Zpost$start
 } else postdur <- 0
   

alpha <- 0.5
colpre <- adjustcolor( "red2", alpha.f = alpha)  
colpost <- adjustcolor( "cyan4", alpha.f = alpha)  
colgrp <- adjustcolor( "gray", alpha.f = alpha)  

basecex <- 2.3

xlim <- range(acou.space[, 1])
# xlim[2] <-  xlim[2] * 1.15

ylim <- range(acou.space[, 2])
# ylim[2] <-  ylim[2] * 1.45

unlink(list.files(pattern = "\\.tiff$"))

tiff("fee%05d.tiff",res = 120, width = 800, height = 800)

if(nrow(Ypre) > 10)
sqpre <- round(seq(1, nrow(Ypre), length.out = stp)[-1]) else sqpre <- 10
  
out <- lapply(sqpre, function(q) {

  plot(x = Y$MDS1, y = Y$MDS2, col = "white", pch = 20, cex = basecex,  xlab = "MDS 1", ylab= "MDS 2", xlim = xlim, ylim = ylim, cex.lab = 1.4) 

  title(paste0(Y$Bird.Name[1], " (", Y$Treatment[1], ")"), cex = 2, line = 2)

  usr <-  par("usr")
  
  rect(par("usr")[1],par("usr")[3],par("usr")[2],par("usr")[4], col = adjustcolor(cols[7], 0.1))

  # progress bar
  sqpb <- seq(usr[1], usr[2], length.out = stp * 2)[-1]
  lines(x = c(usr[1], sqpb[which(sqpre == q)]), y = rep(usr[4] * 1.07, 2), lwd = 7, col = colpre, xpd = TRUE)

  text(x = sqpb[stp/2], y = usr[4] * 1.14, labels = "Pre period", xpd = TRUE)
  
  par(bg = "white")
  polygon(x = acou.space[,1], y = acou.space[,2], col = "white", border = "white", lwd =1)
  
  pp.coords.pre <- as.ppp(as.matrix(Ypre[1:q, c("MDS1", "MDS2")]), usr)
  
  r.coords.pre <- raster(density.ppp(pp.coords.pre))

  msk.pre <- mask(r.coords.pre, acu.sp.sp)

  image(msk.pre, add = T, col=palpre)

  # fix acu space border
  polygon(x = acou.space[,1], y = acou.space[,2], col = adjustcolor("white", 0), border =  "white", lwd =4)
  polygon(x = acou.space[,1], y = acou.space[,2], col = adjustcolor("white", 0), border =  adjustcolor(cols[7], 0.1), lwd =4)

  if (nrow(Ypre) == q) return(msk.pre) else return(NA)
  })

msk.pre <- out[[length(out)]]

if(nrow(Ypost) > 10)
sqpost <- round(seq(1, nrow(Ypost), length.out = stp)[-1]) else sqpost <- 10

out2 <- lapply(sqpost, function(q) {

  plot(x = Y$MDS1, y = Y$MDS2, col = "white", pch = 20, cex = basecex,  xlab = "MDS 1", ylab= "MDS 2", xlim = xlim, ylim = ylim, cex.lab = 1.4) 

  title(paste0(Y$Bird.Name[1], " (", Y$Treatment[1], ")"), cex = 2, line = 2)
   usr <-  par("usr")
  
  rect(par("usr")[1],par("usr")[3],par("usr")[2],par("usr")[4], col = adjustcolor(cols[7], 0.1))
  
  sqpb <- seq(usr[1], usr[2], length.out = stp * 2)[-1]
  lines(x = c(sqpb[stp], sqpb[which(sqpost == q) + stp]), y = rep(usr[4] * 1.07, 2), lwd = 7, col = colpost, xpd = TRUE)

    lines(x = c(usr[1], sqpb[stp]), y = rep(usr[4] * 1.07, 2), lwd = 7, col = colpre, xpd = TRUE)

  text(x = sqpb[stp/2], y = usr[4] * 1.14, labels = "Pre period", xpd = TRUE)

  
  text(x = sqpb[(stp/2) + stp], y = usr[4] * 1.14, labels = "Post period", xpd = TRUE)
  
  par(bg = "white")
  polygon(x = acou.space[,1], y = acou.space[,2], col = "white", border = "white", lwd =1)
  
  pp.coords.post <- as.ppp(as.matrix(Ypost[1:q, c("MDS1", "MDS2")]), usr)
  r.coords.post <- raster(density.ppp(pp.coords.post))

  msk.post <- mask(r.coords.post, acu.sp.sp)

  image(msk.pre, add = T, col=palpre2)

  image(msk.post, add = T, col=palpost)
 
  # fix acu space border
  polygon(x = acou.space[,1], y = acou.space[,2], col = adjustcolor("white", 0), border =  "white", lwd =4)
polygon(x = acou.space[,1], y = acou.space[,2], col = adjustcolor("white", 0), border =  adjustcolor(cols[7], 0.1), lwd =4)
   })


graphics.off()

unlink("fee00001.tiff")

unlink(paste0(Y$Bird.Name[1], ".mp4"))
system(paste0("ffmpeg -framerate 4 -i fee%05d.tiff -c:v libx264 -profile:v high -crf 2 -pix_fmt yuv420p ", Y$Bird.Name[1], ".mp4"))

unlink(paste0(Y$Bird.Name[1], ".gif"))
system(paste0("ffmpeg  -t 20 -ss 00:00:00 -i ", Y$Bird.Name[1],".mp4 ", Y$Bird.Name[1], ".gif"))

unlink(list.files(pattern = "\\.tiff$"))
}

acu_dens_ovlp <- function(Y, acu.sp.sp, mds_prox, usr)
{   
  Ypre <- Y[Y$period == "pre", ]
  Ypost <- Y[Y$period == "post", ]

if(nrow(Ypre) > 0) 
{
pp.coords.pre <- as.ppp(as.matrix(Ypre[, c("MDS1", "MDS2")]), usr)
r.coords.pre <- raster(density.ppp(pp.coords.pre))

msk.pre <- mask(r.coords.pre, acu.sp.sp)

# convert to range 0-1
values(msk.pre) <- (values(msk.pre) - min(values(msk.pre), na.rm = TRUE)) 
values(msk.pre) <- values(msk.pre)/ max(values(msk.pre), na.rm = TRUE)

# keep 95% interval
values(msk.pre)[values(msk.pre) < 0.05] <- NA
}

if(nrow(Ypost) > 0)  
{  
pp.coords.post <- as.ppp(as.matrix(Ypost[, c("MDS1", "MDS2")]), usr)
r.coords.post <- raster(density.ppp(pp.coords.post))

msk.post <- mask(r.coords.post, acu.sp.sp)

# convert to range 0-1
values(msk.post) <- (values(msk.post) - min(values(msk.post), na.rm = TRUE)) 
values(msk.post) <- values(msk.post) / max(values(msk.post), na.rm = TRUE)

# keep 95% interval
values(msk.post)[values(msk.post) < 0.05] <- NA
}

  # test method
  # msk.pre <- msk.post
  # values(msk.pre)[sample(1:length(values(msk.pre)), size = (length(values(msk.pre)))*0.75)] <- 0
  
  
try(values(msk.pre)[!is.na(values(msk.post)) & is.na(values(msk.pre))] <- 0, silent = TRUE)


wgt.ovlp <- try_na(sum(values(msk.post) * (values(msk.post) - abs(values(msk.post) - values(msk.pre))), na.rm = TRUE) / sum(values(msk.post) * values(msk.post), na.rm = TRUE))
return(wgt.ovlp)
}

plot_acu_space <- function(Y, x, mds_prox)
    {
  Ypre <- Y[Y$period == "pre", ]
  Ypost <- Y[Y$period == "post", ]
  Gpost <- mds_prox[mds_prox$period == "post" & mds_prox$Cage == Y$Cage[1] & mds_prox$Bird.Name != bnm[x], ]
# W <- prepost[prepost$POSTpair == Y$POSTpair[1] & prepost$treatment == "Pre", ]
Y <- rbind(Y, Gpost)
  
  #select calls closest to mean acoustic space
 if(nrow(Ypre) > 0)
   {
   mt <- rbind(data.frame(MDS1 = mean(Ypre[,"MDS1"]),MDS2 = mean(Ypre[,"MDS2"])), Ypre[,c("MDS1", "MDS2")])
   
   fspre <- Ypre$sound.files[which.min(as.matrix(dist(mt, upper = T))[-1,1])]
        Zpre <- mtdt[mtdt$sound.files == fspre,, drop = FALSE][1,]
         predur <- Zpre$end - Zpre$start
 } else predur <- 0
        
  if(nrow(Ypost) > 0)
{    
   mt2 <- rbind(data.frame(MDS1 = mean(Ypost[,"MDS1"]),MDS2 = mean(Ypost[,"MDS2"])), Ypost[,c("MDS1", "MDS2")])
    fspost <- Ypost$sound.files[which.min(as.matrix(dist(mt2, upper = T))[-1, 1])]
    Zpost <- mtdt[mtdt$sound.files == fspost, , drop = FALSE][1,]
    postdur <- Zpost$end - Zpost$start
 } else postdur <- 0
   
  if(nrow(Gpost) > 0)
{
         mt3 <- rbind(data.frame(MDS1 = mean(Gpost[,"MDS1"]),MDS2 = mean(Gpost[,"MDS2"])), Gpost[,c("MDS1", "MDS2")])
    fspost2 <- Gpost$sound.files[which.min(as.matrix(dist(mt3, upper = T))[-1, 1])]
    Zpost2 <- mtdt[mtdt$sound.files == fspost2, , drop = FALSE][1, , drop = FALSE]
    postdur2 <- Zpost2$end - Zpost2$start
}  else postdur2 <- 0
 
  #and closest to group centroid
  # mean(Y$MDS1[Y$indiv != x])
  # G <- prepost[prepost$indiv != x,]
  #  fsgroup <- as.character(G$file.sel[which.min(dist(G[,2:3], y = data.frame(mean(G[,2]), mean(G[,3]))))])
  # Gcent <- sel.data[sel.data$sf.sel == fsgroup,]
  # Gdur <- Gcent$end - Gcent$start

  #adjust all to the same duration
maxdur <- max(postdur, predur, postdur2)
    if(maxdur > predur & nrow(Ypre) > 0) 
    {
    Zpre$start <- (Zpre$start - (maxdur-predur)/2)
  Zpre$end <- (Zpre$end + (maxdur-predur)/2)
  } 

  if(maxdur > postdur & nrow(Ypost) > 0) 
    {
    Zpost$start <- (Zpost$start - (maxdur-postdur)/2)
  Zpost$end <- (Zpost$end + (maxdur-postdur)/2)
  } 
  
  if(maxdur > postdur2 & nrow(Zpost2) > 0) 
    {
    Zpost2$start <- (Zpost2$start - (maxdur-postdur)/2)
  Zpost2$end <- (Zpost2$end + (maxdur-postdur)/2)
  } 

   if(nrow(Ypre) > 0)   
    wpre <- read_wave(mtdt, index = which(mtdt$sound.files == Zpre$sound.files), from = Zpre$start, to = Zpre$end)

if(nrow(Ypost) > 0)
      wpost <- read_wave(mtdt, index = which(mtdt$sound.files == Zpost$sound.files), from = Zpost$start, to = Zpost$end)

if(nrow(Zpost2) > 0)
        wpost2 <- read_wave(mtdt, index = which(mtdt$sound.files == Zpost2$sound.files), from = Zpost2$start, to = Zpost2$end)

  # tiff(filename = file.path("/media/twright/E88673588673266A/Budgie analysis/Paper 3 social context FoxP/Acoustics space change with spectros and group centroid", paste0(x, ".tiff")), units = "px", width = 630, height = 480)

  nf <- layout(matrix(c(9, 9, 9, 9, 1, 2, 5, 8, 1, 3, 6, 8, 1, 4, 7, 8), nr = 4, byrow=TRUE), widths=c(0.2, 1.5, 0.11, 3), heights=c(0.2, 1, 1, 1.4))
par(mar = c(0, 0, 0, 0))
plot(1, col = "white", frame.plot = F)
 text(x = 1, y = 1.05, "Frequency (kHz)", srt = 90, cex = 2) 
  par(mar = c(0, 2, 0, 0))
  
  if(nrow(Ypre) > 0)  
  spectro(wpre, f = wpre@samp.rate, wl = wl, ovlp = ovlp, scale = F, osc = F, flim = c(1.1, 5.9), palette = reverse.gray.colors.1, grid = F, axisX = F, tlab = "", cexaxis = 2) else
  plot(1, col = "white", frame.plot = F, axes = F)


if(nrow(Ypost) > 0)  
  spectro(wpost, f = wpost@samp.rate, wl = wl, ovlp = ovlp, scale = F, osc = F, flim = c(1.1, 5.9), palette = reverse.gray.colors.1, grid = F,axisX = F, tlab = "", cexaxis = 2) else
  plot(1, col = "white", frame.plot = F, axes = F)
 

if(nrow(Zpost2) > 0)  
{
  par(mar = c(4, 2, 0, 0))  
  spectro(wpost2, f = wpost2@samp.rate, wl = wl, ovlp = ovlp, scale = F, osc = F, flim = c(1.1, 5.9), palette = reverse.gray.colors.1, grid = F, cexaxis = 2, cexlab = 2)
} else  plot(1, col = "white", frame.plot = F, axes = F)

alpha <- 0.5
colpre <- adjustcolor( "red2", alpha.f = alpha)  
colpost <- adjustcolor( "cyan4", alpha.f = alpha)  
colgrp <- adjustcolor( "gray", alpha.f = alpha)  

# Y$col <- as.character(Y$period)
#   Y$col[Y$period ==  "pre"] <- cols
#   Y$col[Y$period ==  "post"] <- col1 
#     Y$col[Y$period ==  "post"] <- col3 

par(mar = c(0, 0, 0, 0))
plot(1, col = "white", frame.plot = F, axes = F)
rect(par("usr")[1],par("usr")[3],par("usr")[2],par("usr")[4],col = colpre)

par(mar = c(0, 0, 0, 0))
plot(1, col = "white", frame.plot = F, axes = F)
rect(par("usr")[1],par("usr")[3],par("usr")[2],par("usr")[4],col = colpost)

par(mar = c(4, 0, 0, 0))
plot(1, col = "white", frame.plot = F, axes = F, xlab = "")
rect(par("usr")[1],par("usr")[3],par("usr")[2],par("usr")[4],col = colgrp)

par(mar = c(4, 6, 0, 0))  
basecex <- 2.3

#fix plot if points are overlapped by legend
# a <- Y[Y$MDS1 > max(Y$MDS1) - (diff(range(Y$MDS1)) * 0.2) & Y$MDS2 > max(Y$MDS2) - (diff(range(Y$MDS2)) * 0.2),]
# xlim <- c(min(Y$MDS1) - (diff(range(Y$MDS1)) * 0.035), max(Y$MDS1) + (diff(range(Y$MDS1)) * 0.35))
# ylim <- c(min(Y$MDS2) - (diff(range(Y$MDS2)) * 0.035), max(Y$MDS2) + (diff(range(Y$MDS2)) * 0.35))

xlim <- range(acou.space[, 1])
xlim[2] <-  xlim[2] * 1.15

ylim <- range(acou.space[, 2])
ylim[2] <-  ylim[2] * 1.45

# if(nrow(a) > 0)  
plot(x = Y$MDS1, y = Y$MDS2, col = "white", pch = 20, cex = basecex,  xlab = "MDS 1", ylab= "MDS 2", xlim = xlim, ylim = ylim, cex.lab = 2) #else
# plot(x = Y$MDS1, y = Y$MDS2, col = "white", pch = 20, cex = basecex,  xlab = "MDS 1", ylab= "MDS 2", cex.lab = 2)

#plot groups Post calls
# points(x = Y$MDS1[Y$indiv != x], y = Y$MDS2[Y$indiv != x], col = col3, pch = 20, cex = basecex * 1.4)

polygon(x = acou.space[,1], y = acou.space[,2], col = adjustcolor(cols[4], 0.2), border = adjustcolor(cols[4], 0.2), lwd =4)
# plot(acou.space, add = T)
#plot pre and post indiv calls
if(nrow(Zpost2) > 0)  
  points(x = Gpost$MDS1, y = Gpost$MDS2, col = colgrp, pch = 20, cex = basecex* 1.4)

if(nrow(Ypre) > 0)  
points(x = Ypre$MDS1, y = Ypre$MDS2, col = colpre, pch = 20, cex = basecex* 1.4)

if(nrow(Ypost) > 0)  
  points(x = Ypost$MDS1, y = Ypost$MDS2, col = colpost, pch = 20, cex = basecex* 1.4)

if(nrow(Gpost) > 0) 
{
  points(x = Y$MDS1[Y$sound.files == fspost2], y = Y$MDS2[Y$sound.files == fspost2], col = "white", pch = 20, cex = basecex)
    points(x = Y$MDS1[Y$sound.files == fspost2], y = Y$MDS2[Y$sound.files == fspost2], col = colgrp, pch = 20, cex = basecex)}    


if(nrow(Ypre) > 0) 
{points(x = Y$MDS1[Y$sound.files == fspre], y = Y$MDS2[Y$sound.files == fspre], col = "black", pch = 20, cex = basecex * 1.4)
points(x = Y$MDS1[Y$sound.files == fspre], y = Y$MDS2[Y$sound.files == fspre], col = "white", pch = 20, cex = basecex)  
points(x = Y$MDS1[Y$sound.files == fspre], y = Y$MDS2[Y$sound.files == fspre], col = colpre, pch = 20, cex = basecex)
}

if(nrow(Ypost) > 0) 
{   points(x = Y$MDS1[Y$sound.files == fspost], y = Y$MDS2[Y$sound.files == fspost], col = "black", pch = 20, cex = basecex * 1.4)
  points(x = Y$MDS1[Y$sound.files == fspost], y = Y$MDS2[Y$sound.files == fspost], col = "white", pch = 20, cex = basecex)
    points(x = Y$MDS1[Y$sound.files == fspost], y = Y$MDS2[Y$sound.files == fspost], col = colpost, pch = 20, cex = basecex)
}

if(nrow(Gpost) > 0) 
{
  points(x = Y$MDS1[Y$sound.files == fspost2], y = Y$MDS2[Y$sound.files == fspost2], col = "black", pch = 20, cex = basecex)
}    

  #   #plot groups Post calls centroid
  #   mex <- mean(Y$MDS1[Y$indiv != x])
  #   mey <- mean(Y$MDS2[Y$indiv != x])
  #  points(x = mex, y = mey, col = "black", pch = 20, cex = basecex * 1.4)
  # points(x = mex, y = mey, col = "white", pch = 20, cex = basecex)
  #   points(x = mex, y = mey, col = col3, pch = 20, cex = basecex)
  #   
lims <- par("usr")  

legend(lims[2]-((lims[2] - lims[1]) * 0.55), lims[4]-((lims[4] - lims[3]) * 0.002), lty=c(1,1), col=c(colpre, colpost, colgrp, "black"), legend = c("Pre", "Post", "Post group", "Post group centroid"), pch = 20, lwd = 0, pt.cex= basecex * 1.5, cex = basecex * 0.5)
  
par(mar = c(0, 0, 0, 0))
plot(1, col = "white", frame.plot = F, axes = F)
text(1, labels = paste0(bnm[x], " (", Y$Treatment[1], ")"), cex = 2)
# dev.off()
}

overdisp_fun <- function(model) {
  ## number of variance parameters in 
  ##   an n-by-n variance-covariance matrix
  vpars <- function(m) {
    nrow(m)*(nrow(m)+1)/2
  }
  model.df <- sum(sapply(VarCorr(model),vpars))+length(fixef(model))
  rdf <- nrow(model.frame(model))-model.df
  rp <- residuals(model,type="pearson")
  Pearson.chisq <- sum(rp^2)
  prat <- Pearson.chisq/rdf
  pval <- pchisq(Pearson.chisq, df=rdf, lower.tail=FALSE)
  c(chisq=Pearson.chisq,ratio=prat,rdf=rdf,p=pval)
}

```

```{r link metadata, echo=F, eval=T}

pool <- read.csv("Pool sp-dtw-xcorr data 2018.csv", stringsAsFactors = FALSE)

# str(pool)

mtdt <- read.csv("Selections, snr and metadata long term budgie calls 2018.csv", stringsAsFactors = FALSE)

# nrow(mtdt)

ext_sel <- readRDS("budgie call extended selection table may-2018.RDS")

# nrow(ext_sel)

pool$sound.files <- paste(mtdt$sound.files, mtdt$selec, sep = "_")


# pool_mtdt <- merge(pool, mtdt, by = c("sound.files", "selec"))
pool_mtdt <- merge(ext_sel, pool[ , -2], by = "sound.files")

pool_mtdt <- fix_extended_selection_table(pool_mtdt, ext_sel)

# names(pool_mtdt)
pool_mtdt$Cage <- as.character(pool_mtdt$Cage)


pool_mtdt <- pool_mtdt[ , c(which(sapply(pool_mtdt, is.character)), which(sapply(pool_mtdt, is.numeric)))]


pool_mtdt <- pool_mtdt[, !names(pool_mtdt) %in% c("start", "end", "full.path")]


# remove suspicious data 

# 
# table(pool_mtdt$Treatment[pool_mtdt$Bird.Name %in% names(cage.cnt)[cage.cnt  == 2]], pool_mtdt$Bird.Name[pool_mtdt$Bird.Name %in% names(cage.cnt)[cage.cnt  == 2]])
# 
# 
# pool_mtdt$Bird.Name.F <- a(pool_mtdt$Bird.Name, substr(pool_mtdt$Focal.Group, 0, 1))
# 
# mt1 <- as.data.frame.matrix(table(pool_mtdt$Bird.Name[pool_mtdt$Treatment == "Novel"], pool_mtdt$Cage[pool_mtdt$Treatment == "Novel"]))
# 
# mt1 <- as.data.frame.matrix(table(pool_mtdt$Bird.Name[pool_mtdt$Treatment == "Novel"], pool_mtdt$Cage[pool_mtdt$Treatment == "Novel"]))
# 
# mt1[mt1 == 0] <- ""
# 
# mt1
# 
# mt2 <- as.data.frame.matrix(table(pool_mtdt$Bird.Name[pool_mtdt$Treatment == "Stable"], pool_mtdt$Cage[pool_mtdt$Treatment == "Stable"]))
# 
# mt2[mt2 == 0] <- ""
# 
# mt2
# 
# df1 <- as.data.frame.matrix(table(pool_mtdt$Cage, pool_mtdt$Treatment))
# 
# df1[order(df1$Novel, df1$Stable, decreasing = T),]
# 
# length(unique(pool_mtdt$Bird.Name))
# head(pool_mtdt)

```

## Data description
```{r data description, eval = T, echo = F}

print("Number of calls:")
nrow(pool_mtdt)

# print("Data preview:")
# 
# kbl <- kable(pool_mtdt[1:20,], row.names = F, escape = FALSE, format = "html")
# kbl <- kable_styling(kbl, bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = FALSE, font_size = 10)
# scroll_box(kbl, height = "600px", width = "900px",
# box_css = "border: 1px solid #ddd; padding: 5px; ", extra_css = NULL)


print("Experiment:")
unique(pool_mtdt$Experiment)


print("Number of calls per treatment:")


kbl <- kable(as.data.frame(t(as.matrix(table(pool_mtdt$Treatment)))), row.names = F, escape = FALSE, format = "html")

kable_styling(kbl, bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = FALSE, font_size = 16)


kbl <- kable(as.data.frame(t(as.matrix(table(pool_mtdt$Day.Week)))), row.names = F, escape = FALSE, format = "html")

Kbl <- kable_styling(kbl, bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = FALSE, font_size = 16)


print("Number of calls per cage:")

kbl <- kable(as.data.frame(t(as.matrix(table(pool_mtdt$Cage)))), row.names = F, escape = FALSE, format = "html")

kable_styling(kbl, bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = FALSE, font_size = 16)

print("Number of calls per focal group:")

kbl <- kable(as.data.frame(t(as.matrix(table(pool_mtdt$Focal.Group)))), row.names = F, escape = FALSE, format = "html")

kable_styling(kbl, bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = FALSE, font_size = 16)

# print("Number of calls per bird:")
# 
# kbl <- kable(as.data.frame((as.matrix(table(pool_mtdt$Bird.Name)))), row.names = T, col.names = F, escape = FALSE, format = "html")
# 
# kbl <- kable_styling(kbl, bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = FALSE, font_size = 16)
# 
# scroll_box(kbl, height = "600px", box_css = "border: 1px solid #ddd; padding: 5px; ", extra_css = NULL)


# print("Number of calls per bird and treatment:")
# 
# mt <- as.matrix(table(pool_mtdt$Bird.Name, pool_mtdt$Treatment))
# mt[mt == 0] <- ""
# 
# kbl <- kable(mt, row.names = T, escape = FALSE, format = "html")
# 
# kbl <- kable_styling(kbl, bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = FALSE, font_size = 16)
# 
# scroll_box(kbl, height = "600px", box_css = "border: 1px solid #ddd; padding: 5px; ", extra_css = NULL)
# 
# 
# print("Number of calls per cage and bird:")
# 
# mt <- as.matrix(table(pool_mtdt$Bird.Name, pool_mtdt$Cage))
# mt[mt == 0] <- ""
# kbl <- kable(mt, row.names = T, escape = FALSE, format = "html") 
# 
# kbl <- kable_styling(kbl, bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = FALSE, font_size = 16)
# 
# scroll_box(kbl, height = "600px", box_css = "border: 1px solid #ddd; padding: 5px; ", extra_css = NULL)
# 
# 
# print("Number of calls per call type:")
# 
# mt <- as.data.frame.matrix(t(table(pool_mtdt$Call.Type)))
# 
# kbl <- kable(t(mt), row.names = T, col.names = F, escape = FALSE, format = "html") 
# 
# kbl <- kable_styling(kbl, bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = FALSE, font_size = 16)
# 
# scroll_box(kbl, height = "600px", box_css = "border: 1px solid #ddd; padding: 5px; ", extra_css = NULL)
# 
# 
# print("Number of calls per bird and call type:")
# 
# mt <- as.matrix(table(pool_mtdt$Bird.Name, pool_mtdt$Call.Type))
# mt[mt == 0] <- ""
# 
# kbl <- kable(mt, row.names = T, escape = FALSE, format = "html") 
# kbl <- kable_styling(kbl, bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = FALSE, font_size = 16)
# scroll_box(kbl, height = "600px", width = "800px", box_css = "border: 1px solid #ddd; padding: 5px; ", extra_css = NULL)

# 
# print("Individual in 2 cages:")
# 
# mt <- as.matrix(table(pool_mtdt$Bird.Name, pool_mtdt$Cage))
# mt[mt > 0] <- 1
# 
# cage.cnt <- apply(mt, 1, sum)
# 
# names(cage.cnt)[cage.cnt  == 2]
# 
# 
# kbl <- kable(mt, row.names = T, escape = FALSE, format = "html") 
# kbl <- kable_styling(kbl, bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = FALSE, font_size = 16)
# scroll_box(kbl, height = "600px", width = "800px", box_css = "border: 1px solid #ddd; padding: 5px; ", extra_css = NULL)
```

## Correlation matrix

This graph shows the correlation between the parameters used to define the acoustic space. Parameters in which the absolute correlation was higher than 0.95 were removed. Note that random forest is very robust to colinearity. 
```{r curing data, echo=F, eval=T, fig.width= 11, fig.height= 11}

 pool_mtdt$selec <-  as.character(pool_mtdt$selec)
#leave Cerroincognita for testing

dat <- pool_mtdt

pre.dat <- preProcess(dat[, sapply(dat, is.numeric)], method = c("BoxCox", "center", "scale"))

dat[, sapply(dat, is.numeric)] <- predict(pre.dat, dat[, sapply(dat, is.numeric)])

cm <- cor(dat[, sapply(dat, is.numeric)])

#do it without this
high.corr <- findCorrelation(cm, cutoff = .95)

dat <- dat[, -c(high.corr + 2)]

nzv <- nearZeroVar(dat)

if(length(nzv) > 0) dat <- dat[, -c(nzv)]

dat <- dat[, names(dat) != "SNR"]

corrplot::corrplot(cm, bg = "white", outline = F, addgrid.col = NA, order = "AOE")

```

```{r random forest, echo=F, eval=F}

dat$Call.Type <- as.factor(dat$Call.Type)

  rfmodtune <- tuneRF(x = dat[, sapply(dat, is.numeric)], y = dat$Call.Type, importance = F, proximity = F, ntree = 10000, mtryStart = 4, trace = FALSE, doBest = F, plot = F)

rfmod <- randomForest(x = dat[, sapply(dat, is.numeric)], y = dat$Call.Type, importance = T, proximity = T, ntree = 10000, mtry = rfmodtune[which.min(rfmodtune[, 2]), 1])

saveRDS(rfmod, "Random forest model budgie calls 2018.RDS")

```

## Confusion matrix

This is the accuracy of the random forest model at correctly clasifying call types on pre-defined categories based on spectrograph similarity for human observers.

```{r RF results, eval=T, echo=F, fig.width= 11, fig.height= 11}

rfmod <- readRDS("Random forest model budgie calls 2018.RDS")

confusionMatrix(data = rfmod$predicted, reference =  dat$Call.Type)$overall

mat <- rfmod$confusion[,-ncol(rfmod$confusion)]

dev <- as.vector(colSums(mat))
conf.mat <- t(t(mat) / dev)
conf.mat[is.na(conf.mat)] <- 0

col1 <- colorRampPalette(c("#7F0000","red","#FF7F00","yellow","white",
                           "cyan", "#007FFF", "blue","#00007F"))

corrplot(conf.mat, is.corr = FALSE, method = "color", bg = "black")


```

```{r error rate, eval = F, echo=F}
## Test data error rate

mean(rfmod$err.rate[,1])

```

## Variable importance

The relative contribution of each acoustic parameter to the random forest predictive model.Long story short: cross-correlation dereived parameters are the most important ones. 
```{r Variable importance, eval = T, echo=F, include=F}

vimp <- varImpPlot(rfmod, col = "black", pch = 20, pt.cex = 1.5, n.var = length(which(sapply(dat, is.numeric))), main ="Random Forest classification of cultural evolution models")

```

```{r Variable importance 2, eval = T, echo=F, include=T, fig.width = 13, fig.height = 12}

vimp <- as.data.frame(vimp)
vimp$params <- rownames(vimp)

vimp <- vimp[order(vimp$MeanDecreaseAccuracy),]

vimp$params <- as.factor(vimp$params)

vimp$params <- factor(vimp$params, levels = vimp$params[!duplicated(vimp$params)])

theme_set(theme_classic(base_size = 30))
ggplot(vimp, aes(x = MeanDecreaseAccuracy, y = params)) +
  geom_segment(aes(yend = params),  xend = 0, color = "grey50", size = 1) + 
  geom_point(size = 10, col = cols[8]) +
	labs(x = "Mean decrease accuracy (RF)", y = "Predictors") + theme(legend.key.size = unit(2, 'lines')) +  scale_color_discrete(name = "Parameter set") +  scale_fill_discrete(guide = FALSE)

```

```{r RF error, echo=F, eval= F}

fit  <- mds(delta = 1-rfmod$proximity, ndim = 2) 
  
mds_rf_prox <- bootmds(fit, 1-rfmod$proximity, method.dat = "pearson", nrep = 1)

saveRDS(mds_rf_prox, "Random forest MDS budgie calls 2018.RDS")

```

## Random Forest acoustic space (based on bootstrap MDS)

Acoustic space defined by the 2 dimensions of a boostrap multidimensional scale analysis on the random forest proximity [(see here for details](https://stats.stackexchange.com/questions/137358/what-is-meant-by-proximity-in-random-forests)).

```{r, echo=F, eval=T, fig.width= 8, fig.height= 6}

mds_rf_prox <- readRDS("Random forest MDS budgie calls 2018.RDS")

mds_prox <- mds_rf_prox$conf

mds_prox <- data.frame(dat[,!sapply(dat, is.numeric)], MDS1 = mds_prox[,1], MDS2 = mds_prox[,2])

# mds_prox$sound.files[1:300]

ggplot(mds_prox, aes(x = MDS1, y = MDS2, color = Call.Type)) +
  geom_point(size = 2) + 
  # stat_ellipse() +
  # ggtitle("Budgie call acoustic space") +
  theme_classic()

```

```{r acoustic space area change tru time, eval=F, echo=F}

mds_prox$date <- sapply(mds_prox$sound.files, function(x) strsplit(x, " ")[[1]][2])

mds_prox$date <- ifelse(grepl("wav", mds_prox$date), sapply(mds_prox$sound.files, function(x) strsplit(x, " ")[[1]][1]), mds_prox$date)


mds_prox$bird2 <- sapply(mds_prox$sound.files, function(x) strsplit(x, " ")[[1]][1])
mds_prox$bird2 <- sapply(mds_prox$bird2, function(x) strsplit(x, "_")[[1]][1])


mds_prox <- mds_prox[which(mds_prox$bird2 == mds_prox$Bird.Name), ]

mds_prox$date <- gsub("FIE_G", "", mds_prox$date)

mds_prox$date <- as.Date(mds_prox$date, format = "%y-%m-%d")

mds_prox$period <- ifelse(grepl("transfer", mds_prox$Day.Week, ignore.case = TRUE), "post", "pre")

mds_prox <- mds_prox[order(mds_prox$Treatment, mds_prox$Bird.Name, mds_prox$date),]


change.res <- pblapply(unique(mds_prox$Bird.Name), cl = prll, function(x)
  {
  Z <- droplevels(mds_prox[mds_prox$Bird.Name == x, ])
  
    if(nrow(Z) > 1)
    {
    #distance per step
      by <- 6
      stps <- seq(3, nrow(Z) - by, by = by)
      if(stps[length(stps)] < nrow(Z)) stps <- c(stps, nrow(Z)) 
    
      dist1step <- sapply(1:(length(stps) - 1),function(z) dist(Z[c(stps[z], stps[z + 1]), c("MDS1", "MDS2")]))
  
    #area per step
    if(nrow(Z) > 3) {
      area1step <- lapply(1:(length(stps) - 1), function(z)
   {
   area1 <- try(as.numeric(earth.poly(Z[1:stps[z],  c("MDS1", "MDS2")])$area), silent = TRUE) 
  if (class(area1) == "try-error") area1 <- NA

     area2 <- try(as.numeric(earth.poly(Z[1:(stps[z+1]),  c("MDS1", "MDS2")])$area), silent = TRUE) 
  if (class(area2) == "try-error") area2 <- NA

      return(data.frame(abs.area.chng = c(area2 - area1), rel.area.chng = c((area2 - area1)/area1)))
  })

  # area1step[[length(area1step) + 1]] <- data.frame(abs.area.chng = rep(NA, 2), rel.area.chng = rep(NA, 2))

  area1step <- do.call(rbind, area1step)
  
  if (nrow(area1step) > length(dist1step)) dist1step <- c(dist1step, NA)
  if (nrow(area1step) < length(dist1step)) dist1step <- dist1step[-length(dist1step)]

  if (length(stps) > nrow(area1step)) stps <- stps[- length(stps)]
  
  outdf <- data.frame(indiv = Z$Bird.Name[1], day = Z$date[stps], day.week = Z$Day.Week[stps], period = Z$period[stps], dist = dist1step, abs.area.change =  area1step$abs.area.chng, cum.abs.area.change =  cumsum(area1step$abs.area.chng), rel.area.change = area1step$rel.area.chn, cum.rel.area.change = cumsum(area1step$rel.area.chn), n.calls = stps)
  } else
  {  
     if (length(stps) > nrow(area1step)) stps <- stps[- length(stps)]
  outdf <- data.frame(indiv = Z$Bird.Name[1], day = Z$date[stps], day.week = Z$Day.Week[stps], period = Z$period[stps], dist = dist1step, abs.area.change =  NA, cum.abs.area.change = NA, rel.area.change = NA, cum.rel.area.change =NA, n.calls = stps)}
    
   } else   outdf <- data.frame(indiv = Z$Bird.Name[1],  day = NA, day.week = NA, period = NA, dist = NA, abs.area.change =  NA, cum.abs.area.change = NA, rel.area.change = NA, cum.rel.area.change = NA, n.calls = NA)
   
   
return(outdf)
   }
)

change.res <- change.res[which(sapply(change.res, is.data.frame))]

change.res <- do.call(rbind, change.res)

# change.res <- lapply(unique(change.res$indiv), function(x) {
#   X <- change.res[change.res$indiv == x, ]
#   X$n.calls <- seq(3, 100000, by = 3)[1:nrow(X)]
#   return(X)
#   })
# 
# change.res <- do.call(rbind, change.res)

write.csv(change.res, "acoustic space change tru time budgies 2018.csv", row.names = FALSE)


```

## Change in acoustic space area through time

These graphs show, for each individual, the cumulative (pink line) and relative (blue line) change in the area of its acoustic space (convex hull) as new calls are included. Relative change was defined as the change in acoustic space for the last 10 calls at any time. The calls were added in chronological order (as their were produced) and the period in which they occured is shown by the background color. If social group stability plays a role in acoustic "drift", then a bump in the acoustic space would be expected at the start of the "post" period in  birds in  novel social groups.

```{r, eval = T, echo=F, fig.width = 30, fig.height = 100, warning=F, dpi = 40, warning=F, message=F}

change.res <- read.csv( "acoustic space change tru time budgies 2018.csv", stringsAsFactors = FALSE)

change.res$treatment <- sapply(1:nrow(change.res),  function(x) mds_prox$Treatment[mds_prox$Bird.Name == change.res$indiv[x]][1])


change.res1 <- change.res[ , c("indiv", "n.calls", "cum.abs.area.change", "day.week", "period", "treatment")]
names(change.res1)[3] <- "area"
change.res1$area.type <- "Cumulative"

change.res2 <- change.res[ , c("indiv", "n.calls", "abs.area.change", "day.week", "period", "treatment")]
names(change.res2)[3] <- "area"
change.res2$area.type <- "Relative (last 10 calls)"


cr <- rbind(change.res1, change.res2)

crout <- lapply(unique(cr$indiv), function(x) {
  
  X <- cr[cr$indiv == x, ]
  
  X$mx.bx <- NA
    X$min.bx[X$period == "post"] <- X$mx.bx[X$period == "pre"] <- max(X$n.calls[X$period == "pre"], na.rm = TRUE)
  
  X$min.bx[X$period == "pre"] <- 0
  
  X$mx.bx[X$period == "post"] <- Inf

  
X <- X[seq.int(1, nrow(X), length.out = 50), ]

wch.cum <- which(X$area.type == "Cumulative") 
wch.rel <- which(X$area.type != "Cumulative") 
X1 <- X[wch.cum[seq.int(1, length(wch.cum), length.out = 50)], ]
X2 <- X[wch.rel[seq.int(1, length(wch.rel), length.out = 50)], ]

X <- rbind(X1, X2)

  return(X)  
  })

cr <- do.call(rbind, crout)


# remove idnviduals with small samples
cr <- cr[!cr$indiv %in% c("SMO"),]

cr$indiv_treatment <- paste0(cr$indiv, " (", cr$treatment, ")")

cr$indiv_treatment <- factor(cr$indiv_treatment, levels = cr$indiv_treatment[!duplicated(cr$indiv_treatment)][order(cr$treatment[!duplicated(cr$indiv_treatment)])])

cr.per <- cr[!duplicated(cr[, c("indiv","period")]),]


ggplot(cr, aes(x = n.calls, y = area, col = area.type)) +
   geom_rect(aes(xmin =  min.bx, xmax = mx.bx, ymin = 0, ymax = Inf, fill = period), color = NA, size = 0.1, alpha = 0.5, data = cr.per) +
  geom_line(size = 5) +
  geom_point(size = 7, fill = "white", pch = 21) +
  facet_wrap(~indiv_treatment, ncol = 2, scales = "free_x") +
  scale_fill_manual('Period', values = cols[c(5,7)], labels = c("Pre", "Post")) + 
  labs(x = "Number of calls", y = "Acoustic space area") +
  theme_classic(base_size = 40) +
  theme(axis.text.y=element_blank(), axis.ticks.y = element_blank(), legend.position = "top")


```


## Acoustic space plots with spectros

These plots show the acoustic space covered by each individual in 2 ways: 
  - Using dots to show the position of each of its calls in the acoustic space
  - Using kernels to display the density of calls in acoustic space regions 
  
Both plot types are shown consecutively for each individual for comparison.  
  
```{r acoustic space plots with spectros- density plots and density overlap, eval = T, echo = F, out.width=650, dpi = 50}

mtdt <- readRDS("budgie call extended selection table may-2018.RDS")

alpha <- 0.5
wl <- 200
ovlp = 95

mds_rf_prox <- readRDS("Random forest MDS budgie calls 2018.RDS")

mds_prox <- mds_rf_prox$conf

mds_prox <- data.frame(dat[,!sapply(dat, is.numeric)], MDS1 = mds_prox[,1], MDS2 = mds_prox[,2])

mds_prox$date <- sapply(mds_prox$sound.files, function(x) strsplit(x, " ")[[1]][2])

mds_prox$date <- ifelse(grepl("wav", mds_prox$date), sapply(mds_prox$sound.files, function(x) strsplit(x, " ")[[1]][1]), mds_prox$date)

mds_prox$bird2 <- sapply(mds_prox$sound.files, function(x) strsplit(x, " ")[[1]][1])

mds_prox$bird2 <- sapply(mds_prox$bird2, function(x) strsplit(x, "_")[[1]][1])

mds_prox <- mds_prox[which(mds_prox$bird2 == mds_prox$Bird.Name), ]

mds_prox$date <- gsub("FIE_G", "", mds_prox$date)

mds_prox$date <- as.Date(mds_prox$date, format = "%y-%m-%d")

mds_prox$period <- ifelse(grepl("transfer", mds_prox$Day.Week, ignore.case = TRUE), "post", "pre")

mds_prox <- mds_prox[order(mds_prox$Treatment, mds_prox$Bird.Name, mds_prox$date),]

# for acoustic space circle
acou.space <- mcp(SpatialPoints(coords = mds_prox[,9:10]))
acou.space <- attributes(acou.space@polygons[[1]]@Polygons[[1]])$coords

acu.sp.sp <- spPolygons(acou.space[, 1:2])

palpre <-  colorRampPalette(c("white", "red2"))
palpost <-  colorRampPalette(c("white", "cyan4"))

colsn <- 10
palpre2 <- sapply(1: colsn, function(x) adjustcolor(col =  palpre(colsn)[x], alpha = seq(0.2, 0.7, length.out = 20)[x]))
palpre <- sapply(1: colsn, function(x) adjustcolor(col =  palpre(colsn)[x], alpha = seq(0.2, 1, length.out = 20)[x]))
palpost <- sapply(1: colsn, function(x) adjustcolor(col =  palpost(colsn)[x], alpha = seq(0.2, 1, length.out = 20)[x]))

bnm <- mds_prox$Bird.Name[order(mds_prox$Treatment)]

bnm <- bnm[!duplicated(bnm)]

bnm <- bnm[!bnm %in% c("DUV", "PAV")]


# xlim <- range(acou.space[, 1])
# xlim[2] <-  xlim[2] * 1.15
# 
# ylim <- range(acou.space[, 2])
# ylim[2] <-  ylim[2] * 1.45
# 
# plot(x = Y$MDS1, y = Y$MDS2, col = "white", pch = 20, cex = basecex,  xlab = "MDS 1", ylab= "MDS 2", xlim = xlim, ylim = ylim, cex.lab = 2) 
# 
# usr <-  par("usr")

usr <- c(-0.837610,  0.964833, -0.856565,  1.211185)


## show the regions that have been allocated to each plot
plts <- pblapply(1:length(bnm), function(x) {
   Y <- mds_prox[mds_prox$Bird.Name == bnm[x], ]
   tr <- try_na(plot_acu_space(Y, x = x, mds_prox = mds_prox))
   dens.ovlp <- acu_dens_ovlp(Y, acu.sp.sp = acu.sp.sp, mds_prox = mds_prox, usr = usr)
  dens.ovlp <- try_na(round(dens.ovlp, 2))
   tr <- try_na(plot_acu_dens(Y, x = x, mds_prox = mds_prox, ovlp = dens.ovlp))
  }
  )

# save overlaps only

# ovlps <- pbsapply(1:length(bnm), function(x)   acu_dens_ovlp(Y = mds_prox[mds_prox$Bird.Name == bnm[x], ], acu.sp.sp = acu.sp.sp, mds_prox = mds_prox, usr = usr), cl = detectCores() - 1)

# dens.ovlp <- data.frame(bird = bnm, dens.ovlp = ovlps)

# write.csv(dens.ovlp, "post_density_overlap.csv", row.names = FALSE)

```

## Stats

GLMM model:  

$$\ density.post.overlap \sim Treatment + n + Focal.or.group + (1 | Cage)$$

* *Focal.or.Group* predictor was dropped as no effect was detected

```{r stats, eval = T, echo = F}

dens.ovlp <- read.csv("post_density_overlap.csv", stringsAsFactors = FALSE)

names(dens.ovlp)[1] <- "Bird.Name"

mtdt <- readRDS("budgie call extended selection table may-2018.RDS")

dens.ovlp <- merge(dens.ovlp, mtdt, by = "Bird.Name", all.y = F)

dens.ovlp <- dens.ovlp[!duplicated(dens.ovlp$Bird.Name), c("Bird.Name"  ,"dens.ovlp", "Experiment", "Treatment", "Cage", "Focal.Group", "Focal")]

dens.ovlp$n <- sapply(1:nrow(dens.ovlp), function(x) {
  n <- nrow(mtdt[mtdt$Bird.Name == dens.ovlp$Bird.Name[x],])
          if (is.null(n)) n <- NA
  return(n)
          }
)

dens.ovlp <- dens.ovlp[!is.na(dens.ovlp$n), ] 
dens.ovlp$Cage <- as.character(dens.ovlp$Cage)

dens.ovlp.org <- dens.ovlp

pre.dens.ovlp <- dens.ovlp

if (T)
{
  pre.dens.ovlp <- preProcess(dens.ovlp[, sapply(dens.ovlp, is.numeric)], method = c("BoxCox"))

dens.ovlp[, sapply(dens.ovlp, is.numeric)] <- predict(pre.dens.ovlp, dens.ovlp[, sapply(dens.ovlp, is.numeric)])
}

nrow(dens.ovlp)

glob_mod <- lme4::lmer(dens.ovlp ~ Treatment +  n  + Focal.Group + (1 | Cage), data = dens.ovlp, REML = FALSE)


options(na.action = "na.fail")

drg <- dredge(global.model = glob_mod, extra = alist(AIC, BIC, ICOMP, Cp))

drg95 <- drg[1:(max(which(c(0.94, cumsum(drg$weight)) <= 0.95))),]
  
print("Best models")

best_mod <- lmerTest::lmer(dens.ovlp ~ Treatment + n + (1 | Cage), data =  dens.ovlp, REML = FALSE)

print("P values")
summary(best_mod)

if (nrow(drg95) > 1)  
{mod_avg <- model.avg(drg95)
ci <- confint(mod_avg, method = "Wald")} else {mod_avg <- fixef(best_mod)
ci <- confint(best_mod, method = "Wald")
}

ci <- ci[!is.na(ci[,1]), ]
ci <- t(ci)



if (nrow(drg95) > 1)  mod.coeff <- mod_avg$coefficients else
mod.coeff <- t(as.data.frame(fixef(best_mod)))

ci <- rbind(mod.coeff, ci)

ci <- t(as.data.frame(ci))


ci.df <- data.frame(predictor = colnames(mod.coeff), effect = ci[, 1], CI2.5 = ci[, ncol(ci) - 1], CI.97.5 = ci[, ncol(ci)], stringsAsFactors = FALSE)[-1, ]

ci.df$predictor[ci.df$predictor == "TreatmentStable"] <- "Treatment \n (stable-novel)"

ci.df$predictor[ci.df$predictor == "Focal.GroupGroup"] <- "Focal.or.Group"

print("Effect size and confidence intervals")

ggplot(ci.df, aes(y = effect, x = predictor)) + 
  geom_point() +
  geom_errorbar(aes(ymin = CI2.5, ymax = CI.97.5)) +
  labs(y = "Effect size") +
  geom_hline(yintercept = 0, lty = 3) +
 coord_flip()


print("Boxplot")

ggplot(dens.ovlp.org, aes(x = Treatment, y = dens.ovlp)) + 
  geom_boxplot(fill = temp.colors(10)[5]) +
  labs(y = "Post period overlap (%)")

print("test assumptions")

# test assupmptions
qqnorm(resid(best_mod), main="normal qq-plot, residuals", col = topo.colors(1), pch = 20, cex = 2)

qqline(resid(best_mod), col = adjustcolor("red", alpha.f = 0.7))

overdisp(best_mod)

```

```{r gifs, eval = F, echo = F}

mtdt <- readRDS("budgie call extended selection table may-2018.RDS")

alpha <- 0.5
wl <- 200
ovlp = 95

mds_rf_prox <- readRDS("Random forest MDS budgie calls 2018.RDS")

mds_prox <- mds_rf_prox$conf

mds_prox <- data.frame(dat[,!sapply(dat, is.numeric)], MDS1 = mds_prox[,1], MDS2 = mds_prox[,2])

mds_prox$date <- sapply(mds_prox$sound.files, function(x) strsplit(x, " ")[[1]][2])

mds_prox$date <- ifelse(grepl("wav", mds_prox$date), sapply(mds_prox$sound.files, function(x) strsplit(x, " ")[[1]][1]), mds_prox$date)

mds_prox$bird2 <- sapply(mds_prox$sound.files, function(x) strsplit(x, " ")[[1]][1])

mds_prox$bird2 <- sapply(mds_prox$bird2, function(x) strsplit(x, "_")[[1]][1])

mds_prox <- mds_prox[which(mds_prox$bird2 == mds_prox$Bird.Name), ]

mds_prox$date <- gsub("FIE_G", "", mds_prox$date)

mds_prox$date <- as.Date(mds_prox$date, format = "%y-%m-%d")

mds_prox$period <- ifelse(grepl("transfer", mds_prox$Day.Week, ignore.case = TRUE), "post", "pre")

mds_prox <- mds_prox[order(mds_prox$Treatment, mds_prox$Bird.Name, mds_prox$date),]

# for acoustic space circle
acou.space <- mcp(SpatialPoints(coords = mds_prox[,9:10]))
acou.space <- attributes(acou.space@polygons[[1]]@Polygons[[1]])$coords

acu.sp.sp <- spPolygons(acou.space[, 1:2])

palpre <-  colorRampPalette(c("white", "red2"))
palpost <-  colorRampPalette(c("white", "cyan4"))

colsn <- 10
palpre2 <- sapply(1: colsn, function(x) adjustcolor(col =  palpre(colsn)[x], alpha = seq(0.2, 0.7, length.out = 20)[x]))
palpre <- sapply(1: colsn, function(x) adjustcolor(col =  palpre(colsn)[x], alpha = seq(0.2, 1, length.out = 20)[x]))
palpost <- sapply(1: colsn, function(x) adjustcolor(col =  palpost(colsn)[x], alpha = seq(0.2, 1, length.out = 20)[x]))

bnm <- c("SEA", "ZUC", "PLA", "JEW", "TAN", "SWA")

# xlim <- range(acou.space[, 1])
# xlim[2] <-  xlim[2] * 1.15
# 
# ylim <- range(acou.space[, 2])
# ylim[2] <-  ylim[2] * 1.45
# 
# plot(x = Y$MDS1, y = Y$MDS2, col = "white", pch = 20, cex = basecex,  xlab = "MDS 1", ylab= "MDS 2", xlim = xlim, ylim = ylim, cex.lab = 2) 
# 
# usr <-  par("usr")

usr <- c(-0.837610,  0.964833, -0.856565,  1.211185)

# unique(mds_prox$Day.Week)
# 
# mds_prox$week <- 
# as.numeric(gsub("[^\\d]+", "", mds_prox$Day.Week, perl=TRUE))
# 
# unique(gsub("(\\d{1})(?=\\d{2})", "\\1:", mds_prox$week, perl = TRUE))


setwd("~/MEGA/Share/Budgie gifs/")

## show the regions that have been allocated to each plot
plts <- pblapply(1:length(bnm), function(x) {
  
  Y <- mds_prox[mds_prox$Bird.Name == bnm[x], ]
  Y <- Y[order(Y$date), ]
  
  a <-try_na(gif_acu_dens(Y, stp = 20)) 
  
  return(a)  
  })

```


---

---

<font size="4">Session information</font>

```{r session info, echo=F}

sessionInfo()

``0`
